---
title: "MH3510 Assignment 1"
knit: (function(inputFile, encoding) {
      out_dir <- "../output";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir));
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir),
                        output_format = "pdf_document")})
output:
  html_document:
    theme: journal
    toc: true
    toc_float: true
  pdf_document: default
---

# Question

The following data indicate the relationship between the amount of $\beta$-erythroidine in an aqueous solution and the colorimeter reading of the turbidity:

| Concentration | Colorimeter Reading |
|:-------------:|:-------------------:|
| 40            | 69                  |
| 50            | 175                 |
| 60            | 272                 |
| 70            | 335                 |
| 80            | 490                 |
| 90            | 415                 |
| 40            | 72                  |
| 60            | 265                 |
| 80            | 492                 |
| 50            | 180                 |

(a) Fit a simple regression to the data.

(b) Comments on the model adequacy using $\mathcal{F}$ test and $\mathcal{R}^2$ statistics.

# Solution

## Data Preparation

We saved the data in `data/data.csv`. Then we named the columns as $X$ and $y$.

```{r}
data <- read.csv("data/data.csv")
names(data) <- c("X", "y")
head(data)
```

## Simple Linear Regression

We first fit the simple linear regression model using the `lm` function.

```{r}
model <- lm(y ~ X, data = data)
summary(model)
```

We can see that the estimated regression line is

$$\hat{y} = -252.2971 + 8.5290X$$

with $\mathcal{F}\approx 86.829$, $p\approx 1.434\times {10}^{-5}$ and $\mathcal{R}^2\approx 0.9156$.

We can also plot the data and the regression line.

```{r, out.width="70%", fig.align = 'center'}
plot(data$X, data$y, 
     xlab = "Concentration", ylab = "Colorimeter Reading",
     main = "Simple Linear Regression")
abline(model, col = "red")
```

We can also calculate the regression coefficients manually.

We can calculate $\overline{X}$, $\overline{y}$ first.

```{r}
X_bar <- mean(data$X)
y_bar <- mean(data$y)
cat(paste("X_bar =", X_bar, "\ny_bar =", y_bar))
```

Then we calculate $S_{xx}$, $S_{xy}$ and $S_{yy}$ by:

$$
\begin{aligned}
S_{xx} &= \sum_{i=1}^{n} (x_i - \overline{x})^2\\
S_{xy} &= \sum_{i=1}^{n} (x_i - \overline{x})(y_i - \overline{y})\\
S_{yy} &= \sum_{i=1}^{n} (y_i - \overline{y})^2
\end{aligned}
$$

```{r}
S_xx <- sum((data$X - X_bar)^2)
S_xy <- sum((data$X - X_bar) * (data$y - y_bar))
S_yy <- sum((data$y - y_bar)^2)
cat(paste("S_xx =", S_xx, "\nS_xy =", S_xy, "\nS_yy =", S_yy))
```

Then we can calculate the $\hat{\beta}_1$ and $\hat{\beta}_0$ by:

$$
\begin{cases}
\hat{\beta}_1 = \frac{S_{xy}}{S_{xx}}\\
\hat{\beta}_0 = \overline{y} - \hat{\beta}_1\overline{x}
\end{cases}
$$

```{r}
beta_1 <- S_xy / S_xx
beta_0 <- y_bar - beta_1 * X_bar
cat(paste("beta_1 =", beta_1, "\nbeta_0 =", beta_0))
```

So we can write

$$
\hat{y}=-252.2971 + 8.5290X
$$

Which is the same as the result from `lm` function.

## ANOVA Table

It is easy to see the $\mathcal{F}$ statistic and $\mathcal{R}^2$ statistic from the `summary` function. We can also calculate them manually.

We can calculate $\mathrm{SSR}$ and $\mathrm{SSE}$ by:

$$
\begin{aligned}
\mathrm{SSR} &= \sum_{i=1}^{n} (\hat{y}_i - \overline{y})^2\\
\mathrm{SSE} &= \sum_{i=1}^{n} (y_i - \hat{y}_i)^2
\end{aligned}
$$

```{r}
y_hat <- beta_0 + beta_1 * data$X
SSR <- sum((y_hat - y_bar)^2)
SSE <- sum((data$y - y_hat)^2)
cat(paste("SSR =", SSR, "\nSSE =", SSE))
```

And

$$
\begin{aligned}
\mathrm{MS}_{\mathrm{Reg}}&=\mathrm{SSR}\\
s^2&=\frac{1}{n-2}\mathrm{SSE}
\end{aligned}
$$

```{r}
MS_Reg <- SSR
s2 <- SSE / 8
cat(paste("MS_Reg =", MS_Reg, "\ns^2 =", s2))
```

After that, we can obtain the $\mathcal{F}$ statistic by

$$
\mathcal{F} = \frac{\mathrm{MS}_{\mathrm{Reg}}}{s^2}
$$

```{r}
F <- MS_Reg / s2
cat(paste("F =", F))
```

So we can calculate the $p$-value for $\mathcal{H}_0$ is $\beta_1=0$ is

$$
p=\mathbb{P}\left(X, y\mid \beta_1=0\right) = \int_{\mathcal{F}}^{+\infty} F_{1, 8}(f)\mathrm{d}f \approx 1.434\times {10}^{-5}
$$

```{r}
p <- 1 - pf(86.829, 1, 8)
cat(paste("p =", p))
```

The $p$ value is quite small, so we can reject the null hypothesis and say that there is a significant relationship between $X$ (concentration) and $y$ (colorimeter reading).

For $\mathcal{R}^2$ statistic, we can calculate it by

$$
\mathcal{R}^2 = \frac{S_{xy}^2}{S_{xx}S_{yy}}
$$

```{r}
R2 <- S_xy^2 / (S_xx * S_yy)
cat(paste("R2 =", R2))
```

So the $\mathcal{R}^2$ statistic is about $0.9156$. Which means that there is a strong linear relationship between $X$ and $y$.

## Conclusion

For question (a), we have fitted a simple linear regression model to the data by:

$$\hat{y} = -252.2971 + 8.5290X$$

For question (b), we have tested the model adequacy using $\mathcal{F}$ test and $\mathcal{R}^2$ statistics. The $\mathcal{F}$ statistic is $86.829$ and the $p$ value is $1.434\times {10}^{-5}$, which means that there is a significant relationship between $X$ and $y$. The $\mathcal{R}^2$ statistic is $0.9156$, which means that there is a strong linear relationship between $X$ and $y$.
